Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-01 PandaEvolution.tex

Titan’s backfill functionality offers the opportunity to avoid the overhead of
queue wait times without using pilot abstraction. Backfill availability is the
number of worker nodes that cannot be used for a certain amount of time by any
of the jobs already queued on Titan: All queued jobs are either too large or
their walltime is too long. At any point in time, Titan’s Moab scheduler can be
queried for backfill availability. Based upon the result of this query, a job
can be shaped to request no more than the backfill availability. As such, when
submitted this job is usually scheduled immediately, spending almost no time in
the queue.

Compared to pilots, backfill has the disadvantage of limiting the amount of work
nodes that can be requested. Pilots are normal jobs: they can request as many
worker nodes for as much time as a queue can offer. On the contrary, jobs sized
on the basis of backfill information availability \jhanote{should be resource
availability or backfill queue information?}\mtnote{AFAIH, there is no dedicated
queue to backfill} depend on the number of worker nodes that cannot be given to
any other job in the Titan's queue.
% \sergeynote{any other job currently in the Titan's queue?}\mtnote{Better?}.

Usually, backfill availability is a fraction of the total capacity of the queue
but the size of Titan mitigates this limitation. Every year, about 10\% of
Titan's capacity remains unused, corresponding to an average of 30,000 unused
cores. This equals approximately 270M core hours per year, roughly 30\% of the
overall capacity of WLCG.

>>>>

% Titan’s backfill functionality offers the opportunity to avoid the overhead of
% queue wait times without using pilot abstraction. Backfill availability is the
% number of node-hours (or core-hours) that cannot be used for a certain amount
% of time by any of the jobs already queued on Titan: All queued jobs are either
% too large or their walltime is too long. At any point in time, Titan’s Moab
% scheduler can be queried for backfill availability. Based upon the result of
% this query, a job can be shaped to request no more than the backfill
% availability. As such, when submitted this job is usually scheduled
% immediately, spending almost no time in the queue.

Titan’s backfill functionality offers the opportunity to avoid the overhead of
queue wait times without using pilot abstraction. Backfill availability is the
number of node-hours (or core-hours) that  is not utilized by regular jobs
submitted to the queues on Titan. Titan’s backfill functionality offers the
opportunity to avoid the overhead of queue wait times without using pilot
abstraction.  It is important to note that backfill availability is  computed
retrospectively and is determined by Titan's operation team. Backfill
availability represents the theoretical maximum of node-hours that could be
utilized by ATLAS (or any other project) that submit under backfill mode.

Compared to pilots, backfill has the disadvantage of limiting the amount of
work nodes that can be requested. Pilots are normal jobs: they can request as
many worker nodes for as much time as a queue can offer. On the contrary, jobs
sized on the basis of backfill information availability depend on the number
of worker nodes that cannot be given to any other job in the Titan's queue. %
\sergeynote{any other job currently in the Titan's queue?}\mtnote{Better?}.

Backfill availability is typically a small fraction of the total capacity  of
a resource. Notwithstanding,  given the size of Titan this translates into a
substantial capacity. Every year, about 10\% of Titan's capacity remains
unused, corresponding to an average of 30,000 unused cores. This equals
approximately 270M core hours per year, roughly 30\% of the overall capacity
of WLCG.

### 2017-04-01 PandaOverview.tex

PanDA's application model assumes tasks, workloads and workflows. Tasks
represent a set of operations performed on
% a set of events that are
data sets stored in one or more input files. Tasks are decomposed into jobs,
where each job represents the task's set of operations and a partition of the
task's
% events
data set. Since 2005, a certain amount of parallelism has been
progressively introduced for job execution~\cite{crooks2012multi} but, so far,
% no MPI jobs have been considered for production
HEP applications have not required MPI\mtnote{Addresses Alexei's comment. If
this is still too soft, we can write: ``but HEP applications do not require
MPI.''}. Jobs are supposed to be relatively self-contained, capable of setting
up their own execution environment or having a minimal set of common
dependences.

>>>>

PanDA's application model assumes tasks, workloads and workflows. Tasks
represent a set of operations performed on
% a set of events that are
data sets stored in one or more input files. Tasks are decomposed into jobs,
where each job represents the task's set of operations and a partition of the
task's
% events
data set. Since 2005, a certain amount of parallelism has been
progressively introduced for job execution~\cite{crooks2012multi} but, so far,
HEP applications have not required MPI\mtnote{Addresses Alexei's comment. If
this is still too soft, we can write: ``but HEP applications do not require
MPI.''}. Jobs are supposed to be relatively self-contained, capable of setting
up their own execution environment or having a minimal set of common
dependences.
